---
output:
  word_document: default
  html_document: default
---
## Module 2 Assignment 2 - Multiple Linear Regression and Special Issues
## Hackett, Evan

```{r}
library(tidyverse)
library(GGally)
library(MASS)
library(leaps)
```

```{r}
hour <- read_csv("C:/Users/Evan/Desktop/BAN 502/Module 2/Assignment 2/hour.csv")
View(hour)
bike = hour

bike = bike %>% mutate(season = as_factor(as.character(season))) %>%
mutate(season = fct_recode(season,
"Spring" = "1",
"Summer" = "2",
"Fall" = "3",
"Winter" = "4"))

bike = bike %>% mutate(yr = as_factor(as.character(yr))) 
bike = bike %>% mutate(mnth = as_factor(as.character(mnth))) 
bike = bike %>% mutate(hr = as_factor(as.character(hr))) 

bike = bike %>% mutate(holiday = as_factor(as.character(holiday))) %>%
mutate(holiday = fct_recode(holiday,
"NotHoliday" = "0",
"Holiday" = "1"))

bike = bike %>% mutate(workingday = as_factor(as.character(workingday))) %>%
mutate(workingday = fct_recode(workingday,
"NotWorkingDay" = "0",
"WorkingDay" = "1"))

bike = bike %>% mutate(weathersit = as_factor(as.character(weathersit))) %>%
mutate(weathersit = fct_recode(weathersit,
"NoPrecip" = "1",
"Misty" = "2",
"LightPrecip" = "3",
"HeavyPrecip" = "4"))

bike = bike %>% mutate(weekday = as_factor(as.character(weekday))) %>%
mutate(weekday = fct_recode(weekday,
"Sunday" = "0",
"Monday" = "1",
"Tuesday" = "2",
"Wednesday" = "3",
"Thursday" ="4",
"Friday" = "5",
"Saturday" = "6"))
```
The reason that we converted the yr, mnth and hr variables into factors is that if we left them as numbers the numbers, we would only gain one coefficient when running our analysis. By turning these into factors is that the number of coefficients becomes n-1 which will allow us to perform more detailed analysis on the dataset. 

```{r}
ggcorr(bike, label = TRUE)
```

The variable that most correlates with the count variable is the casual variable. 

```{r}
ggplot(bike,aes(x=hr,y=count)) + geom_boxplot()

ggplot(bike,aes(x=yr,y=count)) + geom_boxplot()

ggplot(bike,aes(x=mnth,y=count)) + geom_boxplot()

ggplot(bike,aes(x=holiday,y=count)) + geom_boxplot()

ggplot(bike,aes(x=season,y=count)) + geom_boxplot()

ggplot(bike,aes(x=weekday,y=count)) + geom_boxplot()

ggplot(bike,aes(x=weathersit,y=count)) + geom_boxplot()

ggplot(bike,aes(x=workingday,y=count)) + geom_boxplot()







```

After analyzing the boxplots for our categorical variables I have made the following conclusions about the variables:
hr - as noted in the instructions for the assignment, this variable certainly appears to affect the count.  
yr - this variable also appears to impact the count variable. In the first year the count was around 625, but in the next year it increased upwards to near 1,000.  
mnth - this variable affects count. There is a notable rise in the counts after February, and then starts to decrease around october.  
holiday - this variable impacts the count as well, with the countbeing much higher on days that are not on a holiday. 
Season - this varible does impact the count, but not in the same obcious way as some of the other variables. While the counts are roughly the same, the medians and IQR ranges do change from season to season. 
Weekday - this variable i would argue does not have a significant impact on the count variable. The boxplots are roughly identical across the week. 
Weathersit - this variable plays a significan impact on the count variable. As the weather detoriates the less people are using the bikes.  
Workingday - similar situation to weekday, i would argue this does not play a large part of the count. 

```{r}
bike2 = bike %>%  dplyr::select(-c(instant, dteday, registered, casual))
allmod = lm(count ~., bike2)
summary(allmod)
```


```{r}
emptymod = lm(count ~1, bike2) #use ~1 to build an empty model
summary(emptymod)
```

```{r}
forwardmod = stepAIC(emptymod, direction = "forward",
scope=list(upper=allmod,lower=emptymod),trace = TRUE) #trace = TRUE shows how the model is built (which variables are added)
summary(forwardmod)
```

The forward model includes the following variables: hr, atemp, yr, weathersit, season, mnth, hum, weekday, holiday, windspeed and temp. This roughly aligns with what I predicted would be included after analyzing the boxplots. I argued that the weekday and workingday would not be included, it appears i was roughly correct with the exception of the weekday variable. The non categorical variables make sense when looking at our ggcorr figure, as they did have some correlation. The adjusted rsquared value for the forward stepwise regression is .6854 which is OK.

```{r}
#backward
backmod = stepAIC(allmod, direction = "backward", trace = TRUE)
summary(backmod)
```
The backward model is identical to the forward model. 

The working day variable is accounted for in the other variables such as holiday or weekday. Typically if there is a holiday, it is a non working day, so if it is a non-holiday there is a greater chance of it being a workday. Additionally, for most individuals working days are Monday - Friday, so that ties into the weekday variable. 

```{r}
bike = bike %>% mutate(yr = as.integer(yr)-1)

bike3 = bike %>%  dplyr::select(-c(instant, dteday, registered, casual))
allmod = lm(count ~., bike3)
summary(allmod)
```

```{r}
emptymod = lm(count ~1, bike3) #use ~1 to build an empty model
summary(emptymod)
```

```{r}
forwardmod = stepAIC(emptymod, direction = "forward",
scope=list(upper=allmod,lower=emptymod),trace = TRUE) #trace = TRUE shows how the model is built (which variables are added)
summary(forwardmod)
```
The Model did not change. 
