---
output:
  word_document: default
  html_document: default
---
##Module 4 Assignment 2 
### Hackett, Evan

```{r}
library(tidyverse)
library(caret)
library(ranger)
```


```{r}
Blood <- read_csv("Blood.csv")
View(Blood)

Blood = Blood %>% mutate(DonatedMarch = as_factor(as.character(DonatedMarch))) %>%
mutate(DonatedMarch = fct_recode(DonatedMarch,
"yes" = "1",
"no" = "0"))
```

splitting Blood dataset into training/testing
```{r}
set.seed(1234)
train.rows = createDataPartition(y = Blood$DonatedMarch, p=0.7, list = FALSE) #70% in training
train = Blood[train.rows,] 
test = Blood[-train.rows,]
```

Random forest
```{r}
fit_control = trainControl(method = "cv",  
                           number = 10) #set up 10 fold cross-validation

set.seed(123)  
rf_fit = train(DonatedMarch ~.,    
                 data = train,   
                 method = "ranger",  
                 importance = "permutation", 
                 num.trees = 100,
                 trControl = fit_control)
```

Check out random forest details  
```{r}
varImp(rf_fit)
rf_fit
```

The most important variable is the Total_Donated variable. The Least important variables it the Mnths_Since_Last variable. 

Predictions  
```{r}
predRF = predict(rf_fit, train)
head(predRF)
```

Confusion matrix
```{r}
confusionMatrix(predRF, train$DonatedMarch, positive = "yes")
```

The accuracy of this model is significantly higher than that of the naive model. The accuracy rate of this model is 90.65% which is greater than that of the naive model which has an accuracy rate of 76.15%, this difference is statistically different and we know this as the p value is less than .05. 

Predictions on testing set 
```{r}
predRFTest = predict(rf_fit, test)
head(predRFTest)
```

Confusion matrix
```{r}
confusionMatrix(predRFTest, test$DonatedMarch, positive = "yes")
```
The model does not perform as well on the test data set. The accuracy rate has decreaed to 77.68% which is just marginally better than the naive model of 76.34% and the variance is not statistically significant as the p value is greater than .05. 